Nitric-oxide-feedback to ciliary photoreceptor cells orchestrates UV avoidance

### 2023-01-13 text/Manuscript.Rmd
edit model

The specific expression of NOS in interneurons directly postsynaptic to the cPRCs suggests that NO signalling is involved in the UV response. To test this, we generated two *Platynereis* *NOS* knockout lines with the CRISPR/Cas9 method. We recovered two deletions (NOSΔ11/Δ11 and Δ23/Δ23), both leading to an early stop codon and thus likely representing null alleles (Figure 2A). We could establish a homozygous line for both mutations indicating that NOS is not an essential gene. To quantify UV avoidance, we recorded the trajectories of freely swimming larvae in vertical columns, illuminated laterally from two opposite sides with 395 nm violet light. We also tested phototaxis, by exposing larvae to 480 nm directional and collimated light from the top of the column. As previously shown, wild-type larvae swim downward within 30 sec following non-directional UV/violet light stimulation [@veraszto2018]. In contrast, homozygous *NOS*-mutant larvae showed a strongly diminished UV-avoidance response (Figure 2B and C). This phenotype is similar to the defective UV-avoidance of *c-opsin1* mutant larvae and reveals a requirement for NOS in UV-avoidance behaviour [@veraszto2018]. *NOS*-mutant larvae also showed reduced phototactic behaviour, suggesting a function in the visual eyes mediating phototaxis [@Randel2014a], where *NOS* is also expressed (Figure 1). To test if the observed defects are due to a role of NOS in neuronal signalling or circuit developmental, we also tested larvae exposed to a NOS inhibitor, L-NAME. Larvae incubated for xy min in 0.1 mM or 1 mM L-NAME showed defective UV avoidance, but normal phototaxis. Overall, our results indicate an acute requirement for NOS signalling in UV-avoidance and a possible indirect, developmental role in the visual system.

@@

To characterise the role of NIT-GCs within the cPRC during UV stimulation in more detail, the calcium sensor GCaMP6s were ubiquitously expressed in larvae in which NIT-GCs were knocked down using morpholino to examine changes in calcium levels in the cPRC. Local stimulation of the ciliated part of the cPRC with 405 nm light causes a transient decrease in calcium levels in the cPRC (hyperpolarisation), followed by an increase in calcium levels (depolarisation) when the light is switched off [@veraszto2018]. In NIT-GC1 morphant larvae, 405 nm light stimulation caused a decrease in calcium levels in the cPRC, but the increase in calcium levels that occurred after the light stimulation was stopped was strongly suppressed (Figure 3E). This result suggests that NIT-GC1 mediates depolarising responses after UV stimulation of cPRC and may be involved in extracellular signalling. Interestingly, in NIT-GC2 morphant larvae, an increase in calcium levels occurred immediately after light stimulation (Figure 3F). The different responses of cPRC upon knockdown of NIT-GC1 and NIT-GC2 suggest that there may be different intracellular signalling mechanisms associated with their respective localisations. We focused our analysis on NIT-GC1 because it was localised to cPRC neuroaxons within the neuroplexas region and is involved in extracellular signalling after UV stimulation. It is generally known that depolarisation in the PRC involves elevated cGMPs in the photoreceptors. We therefore used an in vitro system in which cultured cells expressed NIT-GC1 to assess the cGMP activity of NIT-GC1.

@@

**NIT-GC Identification and Phylogenetic Analysis** To identify GPR173 receptors, we obtained a database of vertebrate SREB sequences, including GPR173, GPR85 and GPR27 from (Breton et al. 2021). From these sequences, we produced a Hidden Markov Model (HMM) and used this to mine the 49 species investigated. HMM models were run in HMMR3 with an e-value of 1e−15. The same SREB sequences were used to carry out similarity-based searches using BlastP with an e-value cutoff of 1e−15. We merged these two databases and ran CD-Hit (Fu et al. 2012) to eliminate redundant sequences (at a 99% threshold). To identify the sequences that are closely related to the GPR173 sequences, we ran a cluster-based analysis in CLANS (Frickey and Lupas 2004).The CLANS analysis is available as supplementary fig. S4, Supplementary Material online. To identify clusters, we used the convex-clustering option with 100 jack-knife replicates. The SREB receptors are extremely well conserved and form an easily recognizable cluster. To analyze the phylogeny of SREB receptors, the cluster containing these receptors together with monoaminergic receptors were parsed and used for tree building. We aligned the sequences with MAFFT version 7, with the iterative refinement method E-INS-i. Alignments were trimmed with TrimAl in gappy-out mode (Capella-Gutiérrez et al. 2009). To calculate maximum-likelihood trees, we used IQ-tree2 with the LG+G4 model. To calculate branch support, we ran 1,000 replicates with the aLRT-SH-like and aBayes methods (Minh et al. 2020). The sequences used for the phylogenetic analysis are available in supplementary XX, Supplementary Material online, the trimmed alignment is available in supplementary XX, Supplementary Material online. The raw nexus tree of the SREB receptors is available in supplementary file XX, Supplementary Material online.

**Single-cell analysis** I used Achim et al 2015 for the single-cell data. In fact, in Liz's 2018 paper, she used 107 cells as neurons by removing duplicates from Achim's single-cell data, so I used those cells. Since the raw data were read count data, I normalized them to TPM using Python. After that we converted them to log10. From the sum of the expression levels in 107 cells for each gene, I calculated the percentage in each cell. I chose the appropriate genes for now. For each cell, I identified them with marker genes. After created the data in Python, plotted it using R dot plots.

RPKM? TPM? RPKM calculates the total number of reads per million bp and then divides by the length of each gene, so it is not possible to compare between samples. Instead, TPM first divides by the length of the gene and then divides by the total number of reads per million bases, which allows for more accurate comparisons between samples. In this case I wanted to compare between samples, so I used TPM. In fact, TPM is becoming the mainstream method for many single-cell analyses. The total TPM of each gene between the samples was used to calculate the percentage of expressed genes.

>>>>

The specific expression of NOS in interneurons directly postsynaptic to the cPRCs suggests that NO signalling is involved in the UV response. To test this, we generated two *Platynereis* *NOS* knockout lines with the CRISPR/Cas9 method. We recovered two deletions (NOSΔ11/Δ11 and Δ23/Δ23), both leading to an early stop codon and thus likely representing null alleles (Figure 2A). We could establish a homozygous line for both mutations indicating that NOS is not an essential gene. To quantify UV avoidance, we recorded the trajectories of freely swimming larvae in vertical columns, illuminated laterally from two opposite sides with 395 nm violet light. We also tested phototaxis, by exposing larvae to 480 nm directional and collimated light from the top of the column. As previously shown, wild-type larvae swim downward within 30 sec following non-directional UV/violet light stimulation [@Veraszto2018]. In contrast, homozygous *NOS*-mutant larvae showed a strongly diminished UV-avoidance response (Figure 2B and C). This phenotype is similar to the defective UV-avoidance of *c-opsin1* mutant larvae and reveals a requirement for NOS in UV-avoidance behaviour [@Veraszto2018]. *NOS*-mutant larvae also showed reduced phototactic behaviour, suggesting a function in the visual eyes mediating phototaxis [@Randel2014a], where *NOS* is also expressed (Figure 1). To test if the observed defects are due to a role of NOS in neuronal signalling or circuit developmental, we also tested larvae exposed to a NOS inhibitor, L-NAME. Larvae incubated for xy min in 0.1 mM or 1 mM L-NAME showed defective UV avoidance, but normal phototaxis. Overall, our results indicate an acute requirement for NOS signalling in UV-avoidance and a possible indirect, developmental role in the visual system.

@@

To characterise the role of NIT-GCs within the cPRC during UV stimulation in more detail, the calcium sensor GCaMP6s were ubiquitously expressed in larvae in which NIT-GCs were knocked down using morpholino to examine changes in calcium levels in the cPRC. Local stimulation of the ciliated part of the cPRC with 405 nm light causes a transient decrease in calcium levels in the cPRC (hyperpolarisation), followed by an increase in calcium levels (depolarisation) when the light is switched off [@Veraszto2018]. In NIT-GC1 morphant larvae, 405 nm light stimulation caused a decrease in calcium levels in the cPRC, but the increase in calcium levels that occurred after the light stimulation was stopped was strongly suppressed (Figure 3E). This result suggests that NIT-GC1 mediates depolarising responses after UV stimulation of cPRC and may be involved in extracellular signalling. Interestingly, in NIT-GC2 morphant larvae, an increase in calcium levels occurred immediately after light stimulation (Figure 3F). The different responses of cPRC upon knockdown of NIT-GC1 and NIT-GC2 suggest that there may be different intracellular signalling mechanisms associated with their respective localisations. We focused our analysis on NIT-GC1 because it was localised to cPRC neuroaxons within the neuroplexas region and is involved in extracellular signalling after UV stimulation. It is generally known that depolarisation in the PRC involves elevated cGMPs in the photoreceptors. We therefore used an in vitro system in which cultured cells expressed NIT-GC1 to assess the cGMP activity of NIT-GC1.

@@

**NIT-GC Identification and Phylogenetic Analysis**

To identify GPR173 receptors, we obtained a database of vertebrate SREB sequences, including GPR173, GPR85 and GPR27 from (Breton et al. 2021). From these sequences, we produced a Hidden Markov Model (HMM) and used this to mine the 49 species investigated. HMM models were run in HMMR3 with an e-value of 1e−15. The same SREB sequences were used to carry out similarity-based searches using BlastP with an e-value cutoff of 1e−15. We merged these two databases and ran CD-Hit (Fu et al. 2012) to eliminate redundant sequences (at a 99% threshold). To identify the sequences that are closely related to the GPR173 sequences, we ran a cluster-based analysis in CLANS (Frickey and Lupas 2004).The CLANS analysis is available as supplementary fig. S4, Supplementary Material online. To identify clusters, we used the convex-clustering option with 100 jack-knife replicates. The SREB receptors are extremely well conserved and form an easily recognizable cluster. To analyze the phylogeny of SREB receptors, the cluster containing these receptors together with monoaminergic receptors were parsed and used for tree building. We aligned the sequences with MAFFT version 7, with the iterative refinement method E-INS-i. Alignments were trimmed with TrimAl in gappy-out mode (Capella-Gutiérrez et al. 2009). To calculate maximum-likelihood trees, we used IQ-tree2 with the LG+G4 model. To calculate branch support, we ran 1,000 replicates with the aLRT-SH-like and aBayes methods (Minh et al. 2020). The sequences used for the phylogenetic analysis are available in supplementary XX, Supplementary Material online, the trimmed alignment is available in supplementary XX, Supplementary Material online. The raw nexus tree of the SREB receptors is available in supplementary file XX, Supplementary Material online.

**Single-cell analysis**

I used Achim et al 2015 for the single-cell data. In fact, in Liz's 2018 paper, she used 107 cells as neurons by removing duplicates from Achim's single-cell data, so I used those cells. Since the raw data were read count data, I normalized them to TPM using Python. After that we converted them to log10. From the sum of the expression levels in 107 cells for each gene, I calculated the percentage in each cell. I chose the appropriate genes for now. For each cell, I identified them with marker genes. After created the data in Python, plotted it using R dot plots. RPKM? TPM? RPKM calculates the total number of reads per million bp and then divides by the length of each gene, so it is not possible to compare between samples. Instead, TPM first divides by the length of the gene and then divides by the total number of reads per million bases, which allows for more accurate comparisons between samples. In this case I wanted to compare between samples, so I used TPM. In fact, TPM is becoming the mainstream method for many single-cell analyses. The total TPM of each gene between the samples was used to calculate the percentage of expressed genes.
